---
title: "fishing"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE}

library(tidyverse)
library(tidymodels)
library(gdata)
library(skimr)

theme_set(theme_light())

set.seed(123)

```


#Let's load the data

```{r, echo = F}

fishing <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-08/fishing.csv')

fishing <- fishing %>% filter(year > 1925) # Some data before 1955 have weird behavior


```


# Some data exploratory analysis

```{r}

fishing %>% 
  group_by(year, lake) %>% 
  summarise(year_production = sum(values, na.rm = T)) %>% 
  ggplot(aes(year, year_production)) + 
  geom_line(aes(color = lake)) + 
  scale_y_continuous(labels = label_number_si()) +
  labs(x = NULL, 
       y = "Total Production (in pounds)",
       title = "Total production throughout the years",
       subtitle = "Considering all species together and lakes separate")


```


#It is not feasible to analyze the behaviour of each species on each lake.

```{r}

fishing %>% 
  group_by(year, lake, species) %>% 
  summarise(year_production = sum(values, na.rm = T))
  
```


```{r}

grand_total <- fishing %>%
  group_by(species, year) %>% 
  slice_head(n = 1) %>% 
  select(year, species, grand_total) %>% 
  drop_na() %>% 
  ungroup() %>% 
  group_by(species) %>% 
  mutate(species_max = max(grand_total)) %>% 
         #species = ifelse(species_max <= 10000, "Other", species)) %>% View()
  filter(species_max > 10000) %>% 
  select(year, species, grand_total)

year_total <- grand_total %>% 
  ungroup() %>% 
  group_by(year) %>% 
  summarise(grand_total = sum(grand_total)) %>% 
  mutate(species = "Total")

grand_total <- bind_rows(grand_total, year_total)

grand_total %>% 
  ggplot(aes(year, grand_total)) +
  geom_line(aes(color = species), size = 1) +
  scale_y_continuous(labels = label_number_si()) + 
  labs(x = NULL, 
       y = "Grand total (in pounds)",
       color = NULL,
       title = "Species that had at least one grand total of >10k")


  
```


Let's try to predict the U.S. total production based on the production of Ohio only.

These two plots represent the data used on the model.

```{r}

us_total_production <- fishing %>% 
  filter(region == "U.S. Total") %>% 
  group_by(year) %>% 
  summarise(us_total_production = sum(values, na.rm = T)) 
  
us_total_production %>% 
  ggplot(aes(year, us_total_production)) + 
  geom_line(size = 1) + 
  scale_y_continuous(labels = label_number_si()) +
  labs(x = NULL, 
       y = "Total Production (in pounds)",
       title = "U.S. total production throughout the years",
       subtitle = "Considering all species and all lakes together")


region_production <- fishing %>% 
  group_by(year, region) %>% 
  summarise(region_production = sum(values, na.rm = T)) %>% 
  ungroup() %>% 
  group_by(region) %>% 
  mutate(region_max_production = max(region_production),
         region_min_production = min(region_production)) %>% 
  # This prevents to keep data from regions that did not start fishing activities
  # by the year of 1925.
  filter(region_max_production > 10000, region_min_production > 0) %>% 
  select(-region_min_production, -region_max_production) %>% 
  pivot_wider(names_from = region, values_from = region_production)
# These are the regions that present at least one production of >10k, 
#as shown on a previous plot.

    
region_production %>% 
  ggplot(aes(year, `Ohio (OH)`)) + 
  geom_line(size = 1) +
  scale_y_continuous(labels = label_number_si()) +
  labs(x = NULL, y = "Production (in pounds)",
       title = "Total production of the region of Ohio",
       subtitle = "Considering all species together")

region_production %>% 
  pivot_longer(cols = !starts_with("year"), 
               names_to = "region", values_to = "year_production") %>% 
  ggplot(aes(year, year_production)) + 
  geom_line(aes(color = region), size = 1) +
  scale_y_continuous(labels = label_number_si()) +
  labs(x = NULL, 
       y = "Total Production (in pounds)",
       title = "Total production throughout the years",
       subtitle = "For regions that had, at least, one production of >10k",
       color = NULL)


```

Now, let's define training and testing data.

```{r}


data <- initial_split(region_production)

train_production <- training(data)
test_production <- testing(data)

```


Now, it is possible to fit the model and make the predictions right away.

```{r}

lm_model <- linear_reg() %>% set_engine("lm")

model_fit <- 
  fit(lm_model, `U.S. Total` ~ `Ohio (OH)`, data = train_production)


prediction <- predict(model_fit, new_data = test_production)

```


We can apply some metrics to judge model effectiveness.

```{r}

#Let's bind the real values and the predictions make on the test set.

prediction <- bind_cols(test_production, prediction) %>% 
  ungroup() %>% 
  select(-year)


prediction %>% 
  ggplot(aes(`U.S. Total`, .pred)) +
  geom_abline() +
  geom_point() +
  coord_obs_pred()


pred_metrics <- metric_set(rmse, mae)

prediction %>% 
  ungroup() %>% 
  pred_metrics(truth = `U.S. Total`, estimate = .pred)

```


This model produced a mean absolute error of 13k. Let's try adding more regions to the prediction.


Let's split the data again and fit the model.

```{r}


model_fit <- 
  fit(lm_model, 
      `U.S. Total` ~ `Ohio (OH)` + `Minnesota (MN)` + `Wisconsin (WI)` + 
        `Michigan (MI)` + `MI State Total`, 
      data = train_production)


prediction <- predict(model_fit, new_data = test_production)

```

```{r}
#Let's bind the real values and the predictions make on the test set.

prediction <- bind_cols(test_production, prediction)

prediction %>% 
  ggplot(aes(.pred, `U.S. Total`)) +
  geom_abline() +
  geom_point() +
  coord_obs_pred()


prediction %>% 
  ungroup() %>% 
  pred_metrics(truth = `U.S. Total`, estimate = .pred)

```


Some regions fit the requirement of having least one production of >10k, 
but present data starting only at 1953. Let's try using them on our model.

```{r}

region_production <- fishing %>% 
  filter(year >= 1953) %>% 
  group_by(year, region) %>% 
  summarise(region_production = sum(values, na.rm = T)) %>% 
  ungroup() %>% 
  group_by(region) %>% 
  mutate(region_max_production = max(region_production)) %>% 
  filter(region_max_production > 10000) %>% 
  select(-region_max_production) %>% 
  pivot_wider(names_from = region, values_from = region_production)


region_production %>% 
  pivot_longer(cols = !starts_with("year"), 
               names_to = "region", values_to = "year_production") %>% 
  ggplot(aes(year, year_production)) + 
  geom_line(aes(color = region), size = 1) +
  scale_y_continuous(labels = label_number_si()) +
  labs(x = NULL, 
       y = "Total Production (in pounds)",
       title = "Total production throughout the years",
       subtitle = "For regions that had, at least, one production of >10k",
       color = NULL)

```


```{r}

data <- initial_split(region_production)

train_production <- training(data)
test_production <- testing(data)

```


```{r}
model_fit <- 
  fit(lm_model, 
      `U.S. Total` ~ `Ohio (OH)` + `Mich. Proper (MI)` + `Mich. Proper (WI)` + 
        `Green Bay (MI)` + `Green Bay (WI)`, 
      data = train_production)


prediction <- predict(model_fit, new_data = test_production)
```

```{r}

prediction <- bind_cols(test_production, prediction)

prediction %>% 
  ggplot(aes(.pred, `U.S. Total`)) +
  geom_abline() +
  geom_point() +
  coord_obs_pred()


prediction %>% 
  ungroup() %>% 
  pred_metrics(truth = `U.S. Total`, estimate = .pred)

```

